---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Labels

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

Create one plot on the fuel economy data with customized `title`, `subtitle`, `caption`, `x`, `y`, and `color` labels.

### The Solution

```{r}
mpg |> ggplot() +
    geom_point(
        aes(x = displ, y = cty, color = manufacturer), 
        size = 4) +
    labs(
        title = "Fuel efficiency tend to increase with engine displacement",
        subtitle = "Chevrolet Cars with high engine displacement have significant difference to other manufacturers",
        caption = "Data availible at https://fueleconomy.gov/",
        x = "Engine Displacement (Litres)",
        y = "City Miles per Gallon",
        color = "Car Manufacturer"
    )
```

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Recreate the plot found in this section's exercises using the fuel economy data.

### The Solution

Note the change in the size of the points to better match the original plot and needing labels for both `color` and `shape`.

```{r}
mpg |> ggplot(aes(x = cty, y = hwy)) +
    geom_point(
        aes(color = drv, shape = drv),
        size = 2
        ) +
    labs(
        x = "City MPG",
        y = "Highway MPG",
        color = "Type of drive train",
        shape = "Type of drive train"
    )
```

:::

[source]: <https://r4ds.hadley.nz/communication.html#exercises>
