---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Scales

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

Why doesn't the following code override the default scale?

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)

ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_color_gradient(low = "white", high = "red") +
  coord_fixed()
```

### The Solution

The code above doesn't change the gradient color since `scale_color_gradient()` is for changing the gradient for lines. To get the desired outcome we need to use `scale_fill_gradient()` instead.

```{r}
ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_fill_gradient(low = "white", high = "red") + 
  coord_fixed()
```

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

What is the first argument to every scale? How does it compare to `labs()`?

### The Solution

The first argument to every scale is `name` with a default argument of `waiver()`. `waiver()` is also the default for all the arguments in `lab()`.

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

Change the display of the presidential terms by:

a. Combining the two variants that customize colors and x axis breaks. (Other plots in textbook)
b. Improving the display of the y axis.
c. Labelling each term with the name of the president
d. Adding informative plot labels.
e. Placing breaks every 4 years.

### The Solution



```{r}
#| height: 250px
dat <- presidential |> mutate(id = 33 + row_number())

# Creating vector of 4 year periods starting on (1953
terms <- date("1953-01-20") + years(x = 1:16 * 4)

# combine presidential starts with 4 year periods
p_breaks <- c(terms, presidential$start) |> unique() 

presidential |>
    ggplot(aes(x = start, y = dat$id, color = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = dat$id)) +
    # Combining textbook plots (and add custom breaks)
    scale_x_date(name = NULL, breaks = p_breaks, date_labels = "'%y") +
    scale_color_manual(values = c(Republican = "#E81B23", Democratic = "#00AEF3")) +
    # Adding breaks for each term
    scale_y_continuous(name = "President Num.", breaks = 33:45) +
    # Adding Name Labels
    geom_label(
        data = dat,
        aes(x = end, y = id, label = name)
    )
```

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

First create the following plot. Then, modify the code using `override.aes` to make the legend easier to see.

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(color = cut), alpha = 1/20)
```


### The Solution

To make the legend more clear, I removed the transparancy and increased the size of the points in legend.

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(color = cut), alpha = 1/20) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 4)))
```

:::

[source]: <https://r4ds.hadley.nz/communication.html#exercises-2>
