---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Layout

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(patchwork)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

What happens if you omit the parentheses in the following plot layout. Can you explain why this happens?

```{r}
#| output: false
p1 <- ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  labs(title = "Plot 1")
p2 <- ggplot(mpg, aes(x = drv, y = hwy)) + 
  geom_boxplot() + 
  labs(title = "Plot 2")
p3 <- ggplot(mpg, aes(x = cty, y = hwy)) + 
  geom_point() + 
  labs(title = "Plot 3")

(p1 | p2) / p3
```

### The Solution

```{r}
p1 | p2 / p3
```

Omitting the brackets causes the `|` to be used first. This made the `p1` take the left half while `p2` and `p3` share the right half of the plot.

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Using the three plots from the previous exercise, recreate the following patchwork (check textbook for plot).

### The Solution

To recreate the plot, we to add additional titles to each plot. To do this, we use `plot_annotation()` from the `patchwork` libary.

```{r}
p1 / (p2 | p3) +
  plot_annotation(
    tag_levels = c("A", "B", "C"),
    tag_prefix = "Fig. ",
    tag_suffix = ":"
  )
```

:::

[source]: <https://r4ds.hadley.nz/communication.html#exercises-4>
