---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Data Import

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

What function would you use to read a file where fields were separated with "|"?

### The Solution

I would use `read_delim(delim = '|')`

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Apart from `file`, `skip`, and `comment`, what other arguments do `read_csv()` and `read_tsv` have in common?

### The Solution

  The arguments of `read_csv()` and `read_tsv()` are identical, with the same defaults. 
:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

What are the most important arguments to `read_fwf()`?

### The Solution

The most important arguments to read fixed-width files using `read_fwf()` are the `file` we want to read and `col_positions`, how the entries of the file are separated.

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

To read the following text into a data frame, what argument to `read_csv()` do you need to specify? `"x,y\n1,'a,b'"`

### The Solution

To read the text above, we need to change the default of the argument from `quote = "\""` to `quote = "\'"`.

```{r}
# Adding I() due to readr 2.2.0 changes

read_csv(I("x,y\n1,'a,b'"), quote = "\'")
```

:::

<!-- Next Question -->
## Question 5

::: {.callout-note title='Note' collapse='false'}

### The Problem

Identify what is wrong with each of the following inline CSV files. 

### The Solution

```{r}
#| eval: false
# Observations have more columns than expected
read_csv("a,b\n1,2,3\n4,5,6")

# Rows have more and less entries than columns 
read_csv("a,b,c\n1,2\n1,2,3,4")

# Additional quotation mark 
read_csv("a,b\n\"1")

# I do not see anything wrong with this file
read_csv("a,b\n1,2\na,b")

# File is not separated by commas
read_csv("a;b\n1;3")
```

:::

<!-- Next Question -->
## Question 6

::: {.callout-note title='Note' collapse='false'}

### The Problem

Practice referring to non-syntactic names in the following data frame by:

a. Extracting the variable called 1.
b. Plotting a scatterplot of 1 vs. 2.
c. Creating a new column called 3, which is 2 divided by 1.
d. Renaming the columns to one, two, and three.

### The Solution

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

# (a)
annoying |> select(`1`)

# (b)
annoying |> ggplot(mapping = aes(x = `2`, y = `1`)) +
  geom_point()


# (c)
annoying <- annoying |> mutate(`3` = `2` / `1`) 
annoying

# (d)
annoying |> select(
  one = `1`,
  two = `2`,
  three = `3`,
)

```

:::

[source]: <https://r4ds.hadley.nz/data-import.html#exercises>
