---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Unusual Values

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference in how missing values are handled in histograms and bar charts

### The Solution

```{r}
vals <- c(rep(1, 10), rep(2, 9), NA, rep(3, 10)) |> as.data.frame()
colnames(vals) <- "num"

vals |> ggplot(aes(x = num)) +
    geom_histogram()

vals <- c(rep(1, 10), rep(2, 9), NA, rep(3, 10)) |> as.data.frame()
colnames(vals) <- "num"

vals |> ggplot(aes(x = num)) +
    geom_bar()
```

Both charts remove the missing value from the plot. From my own quick testing and looking at `?geom_bar` and `?geom_histogram`, I could not find any differences in the way both geoms (and stats) handle missing values. When doing further research `geom_bar` may have some options to add a bar to show the number of missing values.

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

What does `na.rm = TRUE` do in `mean()` and `sum()`?

### The Solution

When `na.rm = TRUE` in `mean()`, it will skip missing values and does the calculation as if the missing value does not exist.

```{r}
vals <- c(3,6,9, NA)
mean(vals)
mean(vals, na.rm = TRUE)
```

For `sum()` `na.rm = TRUE` does the same thing as above.

```{r}
sum(vals)
sum(vals, na.rm = TRUE)
```

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

Recreate the frequency plot of `scheduled_dep_time` colored by whether the flight was cancelled or not. Also facet by the `cancelled` variable. Experiment with different values of the `scales` variable in the faceting function to mitigate the effect of more non-cancelled flights than cancelled flights.

### The Solution

There are four different values `scales` can take, all shown below

```{r}
# Tibble from example in this section in textbook
flights13 <- nycflights13::flights |> 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + (sched_min / 60)
  ) 

# default "fixed"
flights13 |>
  ggplot(aes(x = sched_dep_time)) + 
  geom_freqpoly(aes(color = cancelled), binwidth = 1/4) + 
  facet_wrap(~ cancelled) 

# "free"
flights13 |>
  ggplot(aes(x = sched_dep_time)) + 
  geom_freqpoly(aes(color = cancelled), binwidth = 1/4) + 
  facet_wrap(~ cancelled, scale = "free") 

# "free_x"
flights13 |>
  ggplot(aes(x = sched_dep_time)) + 
  geom_freqpoly(aes(color = cancelled), binwidth = 1/4) + 
  facet_wrap(~ cancelled, scale = "free_x") 

# "free_y"
flights13 |>
  ggplot(aes(x = sched_dep_time)) + 
  geom_freqpoly(aes(color = cancelled), binwidth = 1/4) + 
  facet_wrap(~ cancelled, scale = "free_y") 
```

Both the `"free"` and `"free_y"` options both mitigate the issue with the number of observations, since both options change the scale of the axis measuring the number of flights.

:::

[source]: <https://r4ds.hadley.nz/EDA.html#exercises-1>
