---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Aesthetic Mappings

[Full Questions][source]

``` {r libs}
#| output: false
# Loading Packages
library(tidyverse)
```


<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

Create a scatterplot of `hwy` vs. `displ` (from the `mpg` dataset) where the points are pink filled in triangles.

### The Solution

``` {r}
mpg |> ggplot(aes(x = hwy, y = displ)) +
    geom_point(shape = 24, fill = "pink")
```

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Why did the following code not result in a plot with blue points?
```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, color = "blue"))
```

### The Solution

The code above did not create a plot with blue points since we added `color` in the `mapping` argument of `geom_point` instead of the function itself
``` {r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy), color = "blue")
```

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

What does the `stroke` aesthetic do? What shapes does it work with?

### The Solution

The `stroke` aesthetic changes the width of the border of the points. It will work on shapes that have borders to change the width of ().

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`?

### The Solution

If you map color to `displ < 5`, it will split the points into two colors, . It will group the points into colors based on the unique values in the thing that we map into the aesthetic.

``` {r}
mpg |> ggplot() + 
  geom_point(aes(x = displ, y = hwy, color = displ < 5))
```

:::

[source]: <https://r4ds.hadley.nz/layers.html#exercises>
