---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Geometric Objects

[Full Questions][source]

``` {r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

What geom whould you use to draw a line chart? A boxplot? A histogram? An area chart?

### The Solution

The following are the geoms to use for each chart:

- Line chart : `geom_smooth()`
- Boxplot    : `geom_boxplot()`
- Histogram  : `geom_histogram()`, or more generically `geom_freqpoly()`
- Area chart : `geom_area()`, or more generically `geom_ribbon()`

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Earlier in this chapter we used `show.legend` without explaning it:

```{r}
#| eval: false
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(color = drv), show.legend = FALSE)
```

What does `show.legend = FALSE` do here? What happens if you remove it? Why do you think we used it earlier?

### The Solution

It removes the legend of the geom. It was used since the legend of the graphs would block the other graphs.

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

What does the `se` argument to `geom_smooth()` do?

### The Solution

The `se` determine if there is a confidence band around the predicted values (line). Its default value is `TRUE`

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

Recreate the R code necessary to generate the graphs [from the textbook][source]. Note that wherever a categorical variable is used in the plot, it's `drv`.

### The Solution

From left to right, top to bottom:

``` {r}
# Plot 1
mpg |> ggplot(aes(x = displ, y = hwy)) +
    geom_point() + 
    geom_smooth(color = "blue", se = FALSE)

# Plot 2
mpg |> ggplot(aes(x = displ, y = hwy)) +
    geom_point() + 
    geom_smooth(aes(group = drv), se = FALSE)

# Plot 3
mpg |> ggplot(aes(x = displ, y = hwy)) +
    geom_point(aes(color = drv)) + 
    geom_smooth(aes(color = drv), se = FALSE)

# Plot 4
mpg |> ggplot(aes(x = displ, y = hwy)) +
    geom_point(aes(color = drv)) + 
    geom_smooth(color = "blue", se = FALSE)

# Plot 5
mpg |> ggplot(aes(x = displ, y = hwy)) +
    geom_point() + 
    geom_smooth(aes(linetype = drv), color = "blue", se = FALSE)

# Plot 6
mpg |> ggplot(aes(x = displ, y = hwy)) +
    geom_point(color = "white", size = 3) +
    geom_point(aes(color = drv))

```

:::

[source]: <https://r4ds.hadley.nz/layers.html#exercises-1>
