---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Facets

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

What happens if you facet on a continuous variable?

### The Solution

```{r}
mpg |> ggplot(aes(x = hwy)) +
    geom_bar() +
    facet_wrap(~displ)

length(unique(mpg$displ))
```

:::

When you use facet on a continious variable, ggplot creates a graph for each unique value of that variable.

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

What do the empty cells in the plot below with facet_grid(drv ~ cyl) mean? How do the two plots below relate?

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  facet_grid(drv ~ cyl, scales = "free")

ggplot(mpg) + 
  geom_point(aes(x = drv, y = cyl))
```

### The Solution

The empty plots in indicate that there are no observations in the data set with the pair of values associated with `drv` & `cyl`. They relate with the second plot since this plot displays the different pairs of `drv` & `cyl` in the data set.  

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

What plots does the following code make? What does `.` do?

```{r}
#| eval: false
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

### The Solution

The plots make scatterplots split by either `drv` or `cyl` and displayed in a single column or row respectively. The `.` is what is aligning the scatterplots.

```{r}
#| echo: false
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

Compare the plot below with a plot using a color aesthetic mapping. What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) + 
  facet_wrap(~ cyl, nrow = 2)
```

### The Solution

Some advantages to using facets instead of color mapping include:

- More easily seeing the relationship of each group.
- If there are many groups, having too many colors may the plot to become unclear.

Some disadvantages include:

- Not seeing relationships within the whole dataset.
- Some groups may have very few points, which take up space and provides very little information. 

With larger datasets, facets become more appealing since facets will more likely have more observations and plotting all observations into a single plot becomes more cluttered and less useful for inference.

:::

<!-- Next Question -->
## Question 5

::: {.callout-note title='Note' collapse='false'}

### The Problem

Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have `nrow` and `ncol` arguments?

### The Solution

`nrow` and `ncol` determine how many rows or columns there will be when displaying `facet_wrap()`. Other arguments include `drop`, `dir` adn `scales`. `facet_grid()` does not have `nrow` nor `ncol` since the number of rows and columns are determined by the two variables chosen.

:::

<!-- Next Question -->
## Question 6

::: {.callout-note title='Note' collapse='false'}

### The Problem

Which of the following plots makes it easier to compare engine size (`displ`) across cars with different drive trains? What does this say about when to place a faceting variable across rows or columns?

```{r}
ggplot(mpg, aes(x = displ)) + 
  geom_histogram() + 
  facet_grid(drv ~ .)

ggplot(mpg, aes(x = displ)) + 
  geom_histogram() +
  facet_grid(. ~ drv)
```

### The Solution

The first plot is easier for comparision. When determining whether to facet across rows or columns, the dimensions of each plot and which axis care about more across the plots. 

:::

<!-- Next Question -->
## Question 7

::: {.callout-note title='Note' collapse='false'}

### The Problem

Recreate the following plot using `facet_wrap()` instead of `facet_grid()`. How do the positions of the facet labels change?

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

### The Solution

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_wrap(~drv, ncol = 1, strip.position = "right")
```

By default, `facet_wrap` puts labels on `"top"`. Using `strip.position` we can change the position of the label.

:::

[source]: <https://r4ds.hadley.nz/layers.html#exercises-2>
