---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Statistical Transformations

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

What is the default geom associated with `stat_summary()`? How could you rewrite code below to use that geom function instead of the stat function?

```{r}
ggplot(diamonds) + 
  stat_summary(
    aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

### The Solution

The default geom for `stat_summary()` is `"pointrange"`. To use `geom_pointrange()` instead of `stat_summary`, we add `stat = "summary"`.

```{r}
ggplot(diamonds) + 
  geom_pointrange(
    aes(x = cut, y = depth),
    stat = "summary",
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem
 
What does `geom_col()` do?

### The Solution

`geom_col()` makes a bar chart that represents each observation in the aesthetic. It differs from `geom_bar()` since col uses `stat_identity()` instead of `stat_count()`.

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

Most geoms and stats come in pairs that are almost always used in concert. Make a list of all the pairs. What do they have in common?

### The Solution

The stats that we have used that are paired are:

|        Geom         |       Stat        |
| :-----------------: | :---------------: |
| `geom_histogram()`  |   `stat_bin()`    |
|  `geom_density()`   | `stat_density()`  |
|    `geom_bar()`     |  `stat_count()`   |
|   `geom_smooth()`   |  `stat_smooth()`  |
|  `geom_boxplot()`   | `stat_boxplot()`  |
|    `geom_col()`     | `stat_identity()` |
| `geom_pointrange()` | `stat_summary()`  |

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

What variables does `stat_smooth()` compute? What arguments control its behaviour.

### The Solution

`stat_smooth()` computes the following:

- `after_stat(y)` or `after_stat(x)`: estimated values
- `after_stat(ymin)` or `after_stat(xmin)`: lower bound of the confidence band
- `after_stat(ymax)` or `after_stat(xmax)` upper bound of the confidence band
- `after_stat(se)`: Standard error of the estimate

The following arguments control its behaviour:

- `method`
- `formula`
- `n`
- `span`
- `level`
- `method.args`


:::

<!-- Next Question -->
## Question 5

::: {.callout-note title='Note' collapse='false'}

### The Problem

In our proportion bar chart, we needed to set group = 1. Why? In other words, what is the problem with these two graphs?

```{r}
ggplot(diamonds, aes(x = cut, y = after_stat(prop))) + 
  geom_bar()
ggplot(diamonds, aes(x = cut, fill = color, y = after_stat(prop))) + 
  geom_bar()
```

### The Solution

The issue is that without group = 1, the plot is not actually displaying the proportion of the number of observation for each cut.

:::

[source]: <https://r4ds.hadley.nz/layers.html#exercises-3>
