---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Coordinate Systems

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

Turn a stacked bar chart into a pie chart using `coord_polar()`.

### The Solution

I used the bar chart from this section and turned it into a bar chart. 

```{r}
diamonds |> ggplot() + 
  geom_bar(
    mapping = aes(x = clarity, fill = cut), 
    width = 1
  ) + 
  coord_polar()
```

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

What's the difference between `coord_quickmap()` and `coord_map()`?

### The Solution

From looking at `?coord_map`, `coord_quickmap()` uses an approximation that "preserves straight lines" while `coord_map()` uses a 2d projection and may not preserve straight lines. 

**Note:** As of writing this solution, `coord_sf()` is the preferred method to create maps instead of both `coord_quickmap()` and `coord_map()`

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

What does the following plot tell you about the relationship between city and highway mpg? Why is `coord_fixed()` important? What does `geom_abline()` do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

### The Solution

This plot tells us that for all observations the highway mpg is greater than the city mpg. `geom_abline()` creates straight lines on the plot ($y = ax + b$). By default, it creates a line with the function $y = x$, where x and y are plot axes. `coord_fixed()` is important in keeping the units of both axes a fixed ratio, which helps with the clairity of `geom_abline()`. Below is a plot that omits `coord_fixed()` 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline()
```

:::

[source]: <https://r4ds.hadley.nz/layers.html#exercises-5>
