---
execute:
  warnings: false
  message: false
number-sections: false
subtitle: Exercise Solutions
---

# Visualizing Relationships

[Full Questions][source]

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

Run the following lines of code. Which of the two plots is saved as `mpg-plot.png`? Why?

```{r}
#| eval: false
ggplot(mpg, aes(x = class)) +
  geom_bar()
ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
ggsave("mpg-plot.png")
```

### The Solution

The second plot is saved. This is due to it being the most recent plot made, the first plot was overwritten by the second plot.

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

What do you need to change in the code above to save the plot as a PDF instead of a PNG? How could you find out what types of image files would work in `ggsave()`?

### The Solution

To save the plot as a PNG, we would change the string in `ggsave()` to end with `.pdf`.

``` {r}
ggplot(mpg, aes(x = class)) +
  geom_bar()
ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
ggsave("mpg-plot.pdf")
```

Using `?ggsave` shows all the file types that would work.

:::

[source]: <https://r4ds.hadley.nz/data-visualize.html#exercises-3>
