---
title: 1 - Data Visualization
subtitle: Exercise Solutions
---

## [Section 1.2.5](<https://r4ds.hadley.nz/data-visualize.html#exercises>)

``` {r}
#| output: false
# Loading Packages
library(tidyverse)
library(palmerpenguins)
```
 
### 1

::: {.callout-note}
How many row and columns are in `penguins`?
:::

``` {r}
dim(penguins)
```

There are 344 rows and 8 columns in `penguins`.

### 2

::: {.callout-note}
What does the `bill_depth_mm` variable describe?
:::

`bill_depth_mm` represent bill depth of the pegnuins in millimeters, obtained from [CRAN](<https://cran.r-project.org/web/packages/palmerpenguins/refman/palmerpenguins.html>)

### 3

::: {.callout-note}
Make a scatterplot of `bill_depth_mm` vs. `bill_length_mm`.
:::

``` {r}
penguins |> # Following tidyverse style guide by using pipes
    ggplot(mapping = aes(y = bill_depth_mm, x = bill_length_mm)) + 
    geom_point()
```

### 4

::: {.callout-note}
What happens when you make a scatterplot of `species` vs `bill_depth_mm`? What s 
:::

``` {r}
penguins |> 
    ggplot(mapping = aes(y = species, x = bill_length_mm)) + 
    geom_point()
```


### 5

::: {.callout-note}
Why does the following give an error and how would you fix it?
``` {r} 
#| eval: false
ggplot(data = penguins) + 
  geom_point()
```
:::

This plot is missing a value for the `mapping` argument in `ggplot()`.

``` {r}
penguins |>
    ggplot(mapping =  aes(y = bill_depth_mm, x = bill_length_mm)) + 
    geom_point()
```

### 6

::: {.callout-note}
What does the `na.rm` argument do in `geom_point()`? WHat is the default value of the argument. Create a scatter plot where you successfully use this argument set to TRUE.
:::

`na.rm` determines whether points that have missing values create a warning that they were removed. The default value of the argument is `FALSE`. 
``` {r}
penguins |>
    ggplot() + 
    geom_point(
        mapping = aes(y = bill_depth_mm, x = bill_length_mm),
        na.rm = TRUE
        )
```

### 7

::: {.callout-note}
Add the following caption to the plot you made in the previous exercise: "Data came from the palmerpenguins package."
:::

``` {r}
penguins |>
    ggplot() + 
    geom_point(
        mapping = aes(y = bill_depth_mm, x = bill_length_mm),
        na.rm = TRUE
    ) +
    labs(caption = "Data came from the palmerpenguins package.")
```

### 8

::: {.callout-note}
Recreate the following visualization. What aesthetic should `bill_depth_mm` be mapped to? Should it be mapped at the global or geom level?
:::

``` {r}
penguins |>
    ggplot(mapping = aes(y = body_mass_g, x = flipper_length_mm)) + 
    geom_point(mapping = aes(color = bill_depth_mm)) +
    geom_smooth(method = "gam")
```

### 9

::: {.callout-note}
Run this code in your head and predict what the output will look like. Then run the code and check your predictions.
``` {r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)    
```
:::

I predict that the plot will have axes of `body_mass_g` vs `flipper_length_mm`. The points on the plot will be grouped with different by the `island` variable. There will be a linear model for each different island and the standard error of the estimated body mass at each value on the lines will not be visible.

``` {r}
#| output: true
#| echo: false
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)   
```

### 10

::: {.callout-note}
Will these two graphs look different? Why/why not?
``` {r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )
```
:::

These plots will not look different. In the first plot, `geom_point()` and `geom_smooth()` inherit the mapping from `ggplot()`. In the second graph, the same mapping is explicity given to both functions.
