---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Rows

[Full Questions][source]

``` {r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(dplyr)
library(nycflights13)

# Note: warnings and messages are suppressed for all code blocks
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

In a single pipleine for each condition, find all flights that meet the condition:
  
- Had an arrival delay of two or more hours
- Flew to Houston
- Were operated by United, American, or Delta
- Departed in July, August, or September
- Arrived more than two hours late but didn't leave late.
- Were delayed by at least an hour, but made up over 30 minutes in flight.

### The Solution

``` {r}
# Had an arrival delay of two or more hours
flights |> filter(dep_delay >= 120) # dep_delay is in minutes

# Flew to Houston
flights |> filter(dest == "IAH" | dest == "HOU")

# Operated by United, American or Delta
# Note AA: American, DL: Delta, UA: United, from the airlines table 
flights |> filter(carrier %in% c("UA", "AA", "DL"))

# Departed in July, August, or September
flights |> filter(month %in% 7:9)

# Arrived more than two hours late but didn't leave late.
flights |> filter(arr_delay > 120 & dep_delay <= 0)

# Were delayed by at least an hour, but made up over 30 minutes in flight.
# We check second condition by checking the difference
flights |> filter(dep_delay > 60 & (dep_delay - arr_delay) > 30)
```

:::


<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Sort `flights` to find the flights with the longest departure delays. Find the flights that left earliest in the morning.

### The Solution

``` {r}
# Longest departure delays
flights |> arrange(desc(dep_delay))

# Earliest Departures
flights |> arrange(dep_time)
```

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

Sort `flights` to find the fastest flights.

### The Solution

We find the fastest flights by checking the difference in arrival and departure time. Since some flights are overnight, we need to account for rows with `dep_time` > `arr_time` to ensure we do not have negative flight times.

``` {r}
# Add a day (2400 minutes) if dep_time is later than arr_time
# This could produce errors if a flight took more than 24 hours (unrealistic) 
flights |> arrange((arr_time - dep_time) + 2400 * (dep_time > arr_time))
```

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

Was there a flight on every day of 2013?

### The Solution

To check this, we count the number of unique (`month`, `day`) pairs and see if there are 365 unique entries.

``` {r}
flights |> distinct(month, day) |> count()
```

There was a flight on every day of 2013.

:::

<!-- Next Question -->
## Question 5

::: {.callout-note title='Note' collapse='false'}

### The Problem

Which flights traveled the farthest distance? Which traveled the least distance?

### The Solution

We will filter all flights that have `distance` equal to max / min respectively. 

``` {r}
# Maximum flights
tmp <- flights |> filter(distance == max(flights$distance)) 
tmp
tmp |> distinct(origin, dest)


# Minimum flights
tmp <- flights |> filter(distance == min(flights$distance)) 
tmp
tmp |> distinct(origin, dest)
```

JFK to HNL is the longest flight. EWR to LGA is the shortest flight.
:::

<!-- Next Question -->
## Question 6

::: {.callout-note title='Note' collapse='false'}

### The Problem

Does it matter what order you used `filter()` and `arrange()` if you're using both?

### The Solution

It does matter in terms of computational cost. When you arrange first before you filter, we need to sort a larger number of rows compared to if you filter first. Since sorting has a larger computational cost $O(n\cdot\log(n))$ (on average using quicksort) than filtering which has a cost of $O(n)$. Thus we should do filtering first, the operation that is less 'expensive'.

:::

[source]: <https://r4ds.hadley.nz/data-transform.html#exercises>
