---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Columns

[Full Questions][source]

``` {r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(nycflights13)

# Note: warnings and messages are suppressed for all code blocks
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

Compare `dep_time`, `sched_dep_time` and `dep_delay`. How would you expect those three numbers to be related?

### The Solution

The relationship between the three columns should be:

$$dep\_ time - sched\_ dep\_ time = dep\_ delay$$

``` {r}
flights |> select(dep_time, sched_dep_time, dep_delay) |> head()
```

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.

### The Solution

Here are a table of possible ways that obtain these columns using `select()`

| Method         | Pseudocode                                                                |
| -------------- | ------------------------------------------------------------------------- |
| Select columns | `select(dep_time, dep_delay, arr_time, arr_delay`                         |
| Remove columns | `select(-month, !year, ...)`                                              |
| Select Range   | `select(dep_time:arr_delay, -sched_dep_time, -sched_arr_time)`            |
| Remove Ranges  | `select(!<cols_before>, -sched_dep_time, -sched_arr_time, !<cols-after>)` |
:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

What happends if you specify the name of the same variable multiple times in a `select()` call?

### The Solution

``` {r}
flights |> select(tailnum) |> head()
flights |> select(tailnum, tailnum) |> head()
flights |> select(tailnum, tailnum, tailnum) |> head()
```

Repeating the same column multiple times does not change the output of select.

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

What does the `any_of()` function do? Why might it be helpful in conjunction with this vector?

```{r}
variables <- c("year", "month", "day", "dep_delay", "arr_delay")
```

### The Solution

`any_of()` (from `tidyselect`) checks if it first argument is one of the values stored in the second argument. We can use it with `variables` to make our `select()` function call easier to read, reuse, or modify later.

``` {r}
# Now we can change variables if we changes to the select
flights |> select(any_of(variables)) |> head()
```

:::

<!-- Next Question -->
## Question 5

::: {.callout-note title='Note' collapse='false'}

### The Problem

Does the result of running the following code surprise you? How do the select helpers deal with upper and lower case by default? How can you change that default?

```{r}
flights |> select(contains("TIME"))
```

### The Solution

Running the code did surprise me with the amount of columns it returned. By default, select helpers (`contains()`) are case insensitive. To change this, we add the argument `ignore.case = FALSE`.

:::

<!-- Next Question -->
## Question 6

::: {.callout-note title='Note' collapse='false'}

### The Problem

Rename `air_time` to `air_time_min` and move it to the beginning of the data frame.

### The Solution

``` {r}
flights |> 
    rename(air_time_min = air_time) |> 
    relocate(air_time_min, .before = year)
```

:::

<!-- Next Question -->
## Question 7

::: {.callout-note title='Note' collapse='false'}

### The Problem

Why doesn’t the following work, and what does the error mean?

```{r}
#| eval: false
flights |> 
  select(tailnum) |> 
  arrange(arr_delay)
#> Error in `arrange()`:
#> ℹ In argument: `..1 = arr_delay`.
#> Caused by error:
#> ! object 'arr_delay' not found
```

### The Solution

The issue is that during the select function call, we remove all columns except for `tailnum`, so when we try to arrange by `arr_delay`, it does not exist in the table we piped into it. To fix this issue, we just need to swap the order of `select` and `arrange`.

```{r}
flights |> 
  arrange(arr_delay) |>
  select(tailnum)
```

:::

[source]: <https://r4ds.hadley.nz/data-transform.html#exercises-1>
