---
subtitle: Exercise Solutions
number-sections: false
---

# Groups

[Full Questions][source]

``` {r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(nycflights13)
```


<!-- Next Question -->
## Question 1 

::: {.callout-note title='Note' collapse='false'}

### The Problem

Which carrier has the worst average delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not?

### The Solution

``` {r}
flights |> 
    group_by(carrier) |>
    summarize(
        avg_delay = mean(dep_delay, na.rm = TRUE), 
        ) |> 
        arrange(desc(avg_delay)) |>
        head(1)
```

The carrier with the worst average delays is `F9` (Frontier Airlines) with a average delay of 20.2 minutes. Since there are missing values, `na.rm = TRUE` was needed to get proper results. 

It is possible to separate the effects of bad airports versus bad carriers. We can do so by creating a 2-factor ANOVA model (the factors being carriers and airports) and checking if there are any statistically significant increases to delay time given a specific airline or carrier (factor level). However a potential issue could be if the sample size of some (airline, carrier) pairs have a low sample size or no flights, which could cause the model to produce inaccurate results. ***Will Do: ANOVA analysis following this***
:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

Find the flights that are most delayed upon departure to each destination.

### The Solution

The following code produces a table containg the maximum departure delay for each destination airport.

``` {r}
flights |>
    group_by(dest) |>
    relocate(dest, dep_delay) |>
    slice_max(dep_delay, n=1) |>
    arrange(desc(dep_delay))
```
:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

How do delays vary over the dourse of the day? Illustrate your answer with a plot.

### The Solution

First lets look at average departure delays across all departure times: 

``` {r}
dep_delays <- flights |> 
    group_by(dep_time) |>
    summarize(avg_dep_delay = mean(dep_delay))

dep_delays |>
    ggplot() + 
    geom_point(aes(x = dep_time, y = avg_dep_delay), na.rm = TRUE)
```

Next lets look at average arrival delays across all arrival times: 

``` {r}
arr_delays <- flights |> 
    group_by(arr_time) |>
    summarize(avg_arr_delay = mean(arr_delay))

arr_delays |>
    ggplot() + 
    geom_point(aes(x = arr_time, y = avg_arr_delay), na.rm = TRUE)
```

From the two plots, delays are negative (ie flights are early) on average from 06:00 and increase over the day, where on the next day the delays rapidly increase until falling back to early flights. Note for points between times 00:00 - 06:00, the spread is large compared to points between times 06:00 - 23:59. This is due to a extremely small sample size for these times.

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

What happens if you supply a negative `n` to `slice_min()` and friends?

### The Solution

From checking the R help page for the slice functions, having a negative `n` will remove rows equal to the magnitude of `n` (ie -3 will remove 3 rows) from the total number of rows in the table. 

```{r}
temp <- tibble(x = 1:5, y = c('a','b','c','d','e')) 

temp |> slice_min(x, n = -2)
temp |> slice_max(x, n = -4)
```

:::

<!-- Next Question -->
## Question 5

::: {.callout-note title='Note' collapse='false'}

### The Problem

Explain what `count()` does in terms of the dplyr verbs you just learned. What does the `sort` argument to `count()` do?

### The Solution

`count()` will return the number of rows of each group of a table. `sort` is a boolean argument where if `TRUE` will sort the counts in descending order.
:::

<!-- Next Question -->
## Question 6

::: {.callout-note title='Note' collapse='false'}

### The Problem

For the following data, write a prediction for the output, then check if you were correct, and describe what it does.

```{r}
#| echo: false
#| output: false
df <- tibble(
  x = 1:5,
  y = c("a", "b", "a", "a", "b"),
  z = c("K", "K", "L", "L", "K")
)
```

```{r}
#| eval: false 
# Data Frame
df <- tibble(
  x = 1:5,
  y = c("a", "b", "a", "a", "b"),
  z = c("K", "K", "L", "L", "K")
)

# Part a.
df |> group_by(y)

# Part b.
df |> arrange(y)

# Part c.
df |>
  group_by(y) |>
  summarize(mean_x = mean(x))

# Part d.
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

# Part e.
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x), .groups = "drop")

# Part f.
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

df |>
  group_by(y, z) |>
  mutate(mean_x = mean(x))
```

### The Solution


#### a.

The output will be a data frame that starts with all rows that have `y = 'a'`, then will have rows where `y = 'b'` afterwards.
```{r}
df |> group_by(y)
```

My prediction was incorrect, `group_by()` does not alter the order of the rows, it only separates the rows of the data frame into groups. 

#### b.

The output will be a data frame that starts with all rows that have `y = 'a'`, then will have rows where `y = 'b'` afterwards.

```{r}
df |> arrange(y)
```

My prediction was correct, `arrange()` will change the order of the rows of a data frame. This is different to `group_by()` since it does create two groups, as seen in the output.

#### c.


The output will be a table with 2 rows and two columns. The first column will be the unique values of `y` ('a', 'b'). The second column will contain the average of the `x` values associated with the values of `y`.

```{r}
df |>
  group_by(y) |>
  summarize(mean_x = mean(x))
```

My prediction was correct. The pipeline first groups the data frame by `y` values, then it takes that new table and takes the average of all `x` values in each group.

#### d.


The output will be a table with 3 rows and two columns. The first column will be the unique pairs of `y` and `z` values. The second column will contain the average of the `x` values associated with the `(y,z)` pairs.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))
```

My prediction was correct. The pipeline first groups the data frame by unique `(y,z)` pairs, then it takes that new table and takes the average of all `x` values in each group. The message is saying that the data frame that was outputted is no longer grouped by `y` and `z`, it is only group by `y` values. It also says how to hide the message or get the groupings to what it was originally.

#### Part e.

The output will contain the same mean_x values as the previous question, but will not have any group and will not have the message before the output.  

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x), .groups = "drop")
```

My prediction was correct. The pipeline first groups the rows by unique `(y,z)` pairs, then calculates the mean of `x` for each group and ungroups the table. This output is difference since it does not contain any groupings.

:::

#### Part f.

The first output will be the same as the output in Part e. The second output will be a data frame grouped by `(y,z)` pairs that adds a new column that contains the mean of the `x` values in each group, keeping all columns.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

df |>
  group_by(y, z) |>
  mutate(mean_x = mean(x))
```

My prediction for the first output was correct, but my prediction for the second output was incorrect. Since it keeps all rows from the original table, it output `mean_x` for every row in each group, instead of only output for each group. Using `mutate()` also did not produce a message and kept the grouping without the need to add an extra argument. 

[source]: <https://r4ds.hadley.nz/data-transform.html#exercises-2>
