---
subtitle: Exercise Solutions
number-sections: false
---

# First Steps

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(palmerpenguins)

```

<!-- Next Question -->
## Question 1

::: {.callout-note title="Note" collapse="false"}

### The Problem

How many row and columns are in `penguins`?

### The Solution
```{r}
dim(penguins)
```

There are 344 rows and 8 columns in `penguins`.
:::

<!-- Next Question -->
## Question 2

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem


What does the `bill_depth_mm` variable describe?

### The Solution

`bill_depth_mm` represents the bill depth of the pegnuins in millimeters, as stated from [CRAN](<https://cran.r-project.org/web/packages/palmerpenguins/refman/palmerpenguins.html>).

:::


<!-- Next Question -->
## Question 3 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

Make a scatterplot of `bill_depth_mm` vs. `bill_length_mm`.

### The Solution

```{r}
penguins |> # Following tidyverse style guide by using pipes
    ggplot(mapping = aes(y = bill_depth_mm, x = bill_length_mm)) + 
    geom_point() +
    labs(title="bill_depth_mm vs bill_length_mm")
```
:::

<!-- Next Question -->
## Question 4 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

What happens when you make a scatterplot of `species` vs `bill_depth_mm`? What might be a better choice of geom?

### The Solution

```{r}
penguins |> 
    ggplot(mapping = aes(y = species, x = bill_length_mm)) + 
    geom_point() + 
    labs(title="species vs bill_depth_mm")
```

When creating a scatterplot, the points are split into the three different species categories, which is not very helpful when represented by a scatterplto. A boxplot can represent this relationship .

```{r}
penguins |>
    ggplot(mapping = aes(y = species, x = bill_length_mm)) + 
    geom_boxplot()
```
:::


<!-- Next Question -->
## Question 5 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

Why does the following give an error and how would you fix it?
```{r} 
#| eval: false
ggplot(data = penguins) + 
  geom_point()
```

### The Solution

This plot is missing a value for the `mapping` argument in `ggplot()`. Adding this argument fixes the plot.

```{r}
penguins |>
    ggplot(mapping =  aes(y = bill_depth_mm, x = bill_length_mm)) + 
    geom_point()
```

:::


<!-- Next Question -->
## Question 6 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

What does the `na.rm` argument do in `geom_point()`? What is the default value of the argument. Create a scatter plot where you successfully use this argument set to `TRUE`.

### The Solution

`na.rm` determines whether points that have missing values create a warning that they were removed. The default value of the argument is `FALSE`. 

```{r}
penguins |>
    ggplot() + 
    geom_point(
        mapping = aes(y = bill_depth_mm, x = bill_length_mm),
        na.rm = TRUE
        )
```

:::


<!-- Next Question -->
## Question 7 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

Add the following caption to the plot you made in the previous exercise: "Data came from the palmerpenguins package."

### The Solution
```{r}
penguins |>
    ggplot() + 
    geom_point(
        mapping = aes(y = bill_depth_mm, x = bill_length_mm),
        na.rm = TRUE
    ) +
    labs(caption = "Data came from the palmerpenguins package.")
```

:::


<!-- Next Question -->
## Question 8 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

Recreate the following visualization. What aesthetic should `bill_depth_mm` be mapped to? Should it be mapped at the global or geom level?

### The Solution

`bill_depth_mm` should be mapped to `color` in the geom level, otherwise `geom_smooth` would be split by bill depth.

```{r}
penguins |>
    ggplot(mapping = aes(y = body_mass_g, x = flipper_length_mm)) + 
    geom_point(mapping = aes(color = bill_depth_mm), na.rm = TRUE) +
    geom_smooth(method = "gam", na.rm = TRUE)
```

:::


<!-- Next Question -->
## Question 9 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

Run this code in your head and predict what the output will look like. Then run the code and check your predictions.

```{r}
#| eval: false
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)    
```

### The Solution

I predict that the plot will have axes of `body_mass_g` vs `flipper_length_mm`. The points on the plot will be grouped with different by the `island` variable. There will be a linear model for each different island and the standard error of the estimated body mass at each value on the lines will not be visible.

```{r}
#| output: true
#| echo: false
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point(na.rm = TRUE) +
  geom_smooth(se = FALSE, na.rm = TRUE)   
```

I predicted the components correctly, but I was not expecting the lines to overlap as much as they did.

:::


<!-- Next Question -->
## Question 10 

::: {.callout-note title="Note" icon="false" collapse="false"}

### The Problem

Will these two graphs look different? Why/why not?
```{r}
#| eval: false
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )
```

### The Solution

These plots will not look different. In the first plot, `geom_point()` and `geom_smooth()` inherit the mapping from `ggplot()`. In the second graph, the same mapping is explicity given to both functions.



```{r}
#| echo: false
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(na.rm = TRUE) +
  geom_smooth(na.rm = TRUE) +
  labs(title = "Plot 1")

ggplot() +
  geom_point(
    data = penguins, 
    mapping = aes(x = flipper_length_mm, y = body_mass_g),
    na.rm = TRUE
  ) +
  geom_smooth(
    data = penguins, 
    mapping = aes(x = flipper_length_mm, y = body_mass_g),
    na.rm = TRUE
  ) +
  labs(title = "Plot 2")
```

:::

[source]: <https://r4ds.hadley.nz/data-visualize.html#exercises>
