---
subtitle: Exercise Solutions
number-sections: false
---

# Visualizing Distributions

[Full Questions][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(palmerpenguins)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title="Note" collapse="false"}

### The Problem

Make a bar plot of `species` of `penguins`, where you assign `species` to the `y` aesthetic. How is this plot different?

### The Solution
```{r}
penguins |>
    ggplot(aes(y = species)) +
    geom_bar()
```

This plot has horizontal bars instead of vertical bars.

:::


<!-- Next Question -->
## Question 2

::: {.callout-note title="Note" collapse="false"}

### The Problem

How are the following two plots different? Which aesthetic, `color` or `fill`, is more useful for changing the color of bars?

```{r}
ggplot(penguins, aes(x = species)) +
  geom_bar(color = "red")

ggplot(penguins, aes(x = species)) +
  geom_bar(fill = "red")
```

### The Solution

The `color` aesthetic changes the border of the bars while `fill` changes the color of the whole bar. `fill` is the more useful option when changing bar colours.

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title="Note" collapse="false"}

### The Problem

What does the `bins` argument in `geom_histogram()` do?

### The Solution

The `bins` argument in `geom_histogram()` determines the exact number of bins, the default number is 30.

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title="Note" collapse="false"}

### The Problem

Make a histogram of the carat variable in the diamonds dataset. Experiment with different binwidths. What binwidth reveals the most interesting patterns?

### The Solution

As we decrease the binwidth in the histograms, we can see decreasing the binwidth shows peaks that were hidden from the larger binwidths. From the different bandwidths tested, `binwidth = 0.1` produced the best results. A smaller binwidth caused the historgram to be visually unclear.

```{r}
# Quick look of the carat variable
summary(diamonds$carat)

diamonds |>
    ggplot(aes(x = carat)) +
    geom_histogram(binwidth = 1)

diamonds |>
    ggplot(aes(x = carat)) +
    geom_histogram(binwidth = 0.5)

diamonds |>
    ggplot(aes(x = carat)) +
    geom_histogram(binwidth = 0.25)

diamonds |>
    ggplot(aes(x = carat)) +
    geom_histogram(binwidth = 0.1)

diamonds |>
    ggplot(aes(x = carat)) +
    geom_histogram(binwidth = 0.01)

# Extremely small "binwidth"
diamonds |>
    ggplot(aes(x = carat)) +
    geom_density()
```

:::

[source]: <https://r4ds.hadley.nz/data-visualize.html#exercises-1>
