---
subtitle: Exercise Solutions
execute:
  warnings: false
  message: false
number-sections: false
---

# Logical Vectors

[Textbook Section][source]

```{r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(nycflights13)
library(palmerpenguins)
```

## 12.2 Comparisons

<!-- Next Question -->
### Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

How does `dplyr::near()` work? Type `near` to see the source code. Is `sqrt(2)^2` near 2?

### The Solution

```{r}
near
```

`near` functions by checking if the absolute difference between the two values is less than a tolerance, where the default is `.Machine$double.eps^0.5` = `r .Machine$double.eps^0.5`, 

```{r}
near(2, sqrt(2)^2)
```

`sqrt(2)^2` is near 2.

:::

<!-- Next Question -->
### Question 2 

::: {.callout-note title='Note' collapse='false'}

### The Problem

Use `mutate()`, `is.na()`, and `count()` together to describe how the missing values in `dep_time`, `sched_dep_time`, and `dep_delay` are connected 

### The Solution

```{r}
flights |>
    mutate(
        miss_dep_time = is.na(dep_time),
        miss_sched_dep = is.na(sched_dep_time),
        miss_dep_delay = is.na(dep_delay)
    ) |> 
    count(miss_dep_time, miss_sched_dep, miss_dep_delay)
```

From the tibble of counts above, there are no missing values in the `miss_sched_dep` column. Every row that has missing value in the `dep_time` column also has a missing value in `dep_delay` and vice versa. 

:::

## 12.3 Boolean Algebra

<!-- Next Question -->
### Question 1 

::: {.callout-note title='Note' collapse='false'}

### The Problem

Find all flights where `arr_delay`  is missing but `dep_delay` is not. Find all flights where neither `arr_time` nor `sched_arr_time` are missing, but `arr_delay` is.

### The Solution

Note the statement A nor B is equivalent to (not A) and (not B).

```{r}

f1 <- flights |> filter(is.na(arr_delay) & !is.na(dep_delay))
f2 <- flights |> 
    filter(
        !(is.na(arr_time)) & 
        !(is.na(sched_arr_time)) & 
        is.na(arr_delay)
    )
count(f1)
count(f2)
```

:::

There are `r count(f1)` flights that follow the first statement and `r count(f2)` flights that follow the second statement.

<!-- Next Question -->
### Question 2 

::: {.callout-note title='Note' collapse='false'}

### The Problem

How many flights have a missing `dep_time`? What other variables are missing in these rows? What might these rows represent?

### The Solution

```{r}
f3 <- flights |> filter(is.na(dep_time))

```

There are `r count(f3)` flights that have a missing `dep_time`. The columns `dep_delay` and `arr_time` are are also missing in these rows. These rows could represent flights that did not depart, would mean the flights would not have a departure delay or an arrival time.

:::

<!-- Next Question -->
### Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

Assuming that a missing `dep_time` implies that a flights is cancelled, look at the number of cancelled flights per day. Is there a pattern? Is there a connection between the proportion of cancelled flights and the average delay of non-cancelled flights?

### The Solution

```{r}
f4 <- mutate(f3, date = paste(year, month, day, sep = "-")) |>
    group_by(date) |>
    count() |>
    arrange(desc(n))
head(f4)

f5 <- flights |> 
    filter(!is.na(dep_time)) |>
    mutate(date = paste(year, month, day, sep = "-")) |>
    group_by(date) |>
    summarize(
        avg_dep_delay = mean(dep_delay),
        # Get number of non-cancelled flights
        num_flown_flights = sum(dep_delay == dep_delay) 
        )

left_join(f4,f5) |>
    mutate(prop_cancelled = n / (n + num_flown_flights)) |>
    ggplot(aes(x = avg_dep_delay, y = prop_cancelled)) + 
    geom_point() + 
    geom_smooth()
```

There are a high number of flights cancelled during dates that are typical for severe weather conditions (heavy snow and rain). From the plot above there seems to be a positive relationship between the average departure delay and the proportion of cancelled flights.

:::

## 12.4 Summaries

<!-- Next Question -->
### Question 1  

::: {.callout-note title='Note' collapse='false'}

### The Problem

What will `sum(is.na(x))` tell you? How about `mean(is.na(x))`?

### The Solution

`sum(is.na(x))` will tell you the total number of elements of `x` that are missing. `mean(is.na(x))` will tell you the proportion of missing values in all elements`x`.

```{r}
# Note x contains 4 NAs and 10 elements in total.
x <- c(NA, 1, 2, 3, NA, NA, 4, 5, NA, -10)

sum(is.na(x))

mean(is.na(x))
```

:::

<!-- Next Question -->
### Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

What does `prod()` return when applied to a logical vector? What logical summary function is it equivalent to? What does `min()` return when applied to a logical vector? What logical summary function is it equivalent to? Read the documentation and perform a few experiments.

### The Solution

When given a logical vector, `prod()` and `min()` handles `TRUE` as if it were `1` and `FALSE` as if it were `1`. This causes both to be equivalent to the function `all()`, since both will only equal `1` (ie `TRUE`) if all elements in the vector are `TRUE`, 0 (ie `FALSE`) otherwise.
```{r}
# prod
prod(c(TRUE, TRUE))
prod(c(FALSE, TRUE))
prod(c(FALSE, FALSE))

# min
min(c(TRUE, TRUE))
min(c(FALSE, TRUE))
min(c(FALSE, FALSE))

# all
all(c(TRUE, TRUE))
all(c(FALSE, TRUE))
all(c(FALSE, FALSE))
```

:::
 
## 12.5 Conditional Transformations

<!-- Next Question -->
### Question 1 

::: {.callout-note title='Note' collapse='false'}

### The Problem

A number is even if it's divisible by two, which in R you can find out with `x %% 2 == 0`. Use this fact and if_else() to determine whether each number between 0 and 20 is even or odd.

### The Solution

```{r}
x <- c(1, 2, 3, 5, 7, 11, 13)
if_else(x %% 2 == 0, "even", "odd")
```

:::

<!-- Next Question -->
### Question 2 

::: {.callout-note title='Note' collapse='false'}

### The Problem

Given a vector of days like "x <- c("Monday", "Saturday", "Wednesday")", use an if_else() statement to label them as weekends or weekdays.

### The Solution

Assume all days are capitalized.

```{r}
days <- c(
        "Monday", "Tuesday", "Wednesday", 
        "Thusday", "Friday", "Saturday", "Sunday" )

if_else(days %in% c("Sunday", "Saturday"), "Weekend", "Weekday")
```

:::

<!-- Next Question -->
### Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

Use `if_else()` to compute the absolute value of a numeric vector called `x`.

### The Solution

```{r}
x <- c(-1, 10, -2.4, 0, 0, 11)

if_else(x > 0, x, -x)
```

:::

<!-- Next Question -->
### Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

When a `case_when()` statement that uses the `month` and `day` columns from `flights` to label a selection of important US holidays (e.g., New Years Day, 4th of July, Thanksgiving, and Christmas). First create a logical column that is either `TRUE` or `FALSE`, and then create a character column that either gives the name of the holiday or is `NA`.

### The Solution

```{r}
flights |> 
    mutate(
        is_holiday = case_when(
        month == 1 & day == 1 ~ TRUE,   # New Years Day
        month == 7 & day == 4 ~ TRUE,   # 4th of July
        month == 12 & day == 25 ~ TRUE, # Christmas Day
        month == 11 & day == 26 ~ TRUE, # 2026 US Thanksgiving
        .default = FALSE
        ),
        holiday = case_when(
        month == 1 & day == 1 ~ "New Years Day",
        month == 7 & day == 4 ~ "4th of July",
        month == 12 & day == 25 ~ "Christmas Day",
        month == 11 & day == 26 ~ "2026 US Thanksgiving",
        .default = NA
        ),
    ) |>
    relocate(is_holiday, holiday) |>
    group_by(holiday) |>
    count()
```

:::

[source]: <https://r4ds.hadley.nz/logicals.html>
