---
execute:
  warnings: false
  message: false
number-sections: false
subtitle: Exercise Solutions
---

# Workflow: Basics

[Full Questions][source]

``` {r libs}
#| output: false
# Loading Packages
library(tidyverse)
library(palmerpenguins)
```

<!-- Next Question -->
## Question 1

::: {.callout-note title='Note' collapse='false'}

### The Problem

Why does this code not work?

``` {r}
#| eval: false
my_variable <- 10
my_varıable
#> Error:
#> ! object 'my_varıable' not found
```

### The Solution

The variable names are not the same.

:::

<!-- Next Question -->
## Question 2

::: {.callout-note title='Note' collapse='false'}

### The Problem

```{r}
#| eval: false
libary(todyverse)

ggplot(dTA = mpg) + 
  geom_point(maping = aes(x = displ y = hwy)) +
  geom_smooth(method = "lm)
```

### The Solution

``` {r}
# dTA -> data, move aes, fix mapping typo, add comma
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + #
  geom_smooth(method = "lm") # add ending quote
```

:::

<!-- Next Question -->
## Question 3

::: {.callout-note title='Note' collapse='false'}

### The Problem

Press Alt + Shift + K. What happens? How can you get to the same place using the menus? 

### The Solution

In RStudio (I am using VSCode) pressing Alt + Shift + K shows a reference for keyboard shortcuts. To get to this using only menus, go from Help to Keyboard Shortcuts Help.

:::

<!-- Next Question -->
## Question 4

::: {.callout-note title='Note' collapse='false'}

### The Problem

Run the following lines of code. Which of the two plots is saved as mpg-plot.png? Why?

```{r}
#| eval: false
my_bar_plot <- ggplot(mpg, aes(x = class)) +
  geom_bar()
my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
ggsave(filename = "mpg-plot.png", plot = my_bar_plot)
```

### The Solution

`my_bar_plot` is saved since it is explicitly added as an argument to `ggsave()`

:::

[source]: <https://r4ds.hadley.nz/workflow-basics.html#exercises>
